---
title: "Interacting with API's - Example with the Open Movie Database API"
author: "Grace Holliday"
date: "2023-06-11"
output: github_document
  toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('httr')
library('jsonlite')
library('tidyverse')
library('imager')
library('ggplot2')

```

This document is a vignette designed to introduce retrieving and manipulating
data from an API (Application Programming Interface). I will be interacting 
with the Open Movie Database API (OMDb).  I will build several functions 
comprised of a six endpoints and demonstrate some basic data manipulation. 

Endpoints include: Highest Grossing Movie, Highest Average Movie Rating,
Runtime by Genre, Viewing Movie Posters, Oldest Movies, and Movies by Genre
and Rating (PG, PG-13, etc.)

### Requirements

Packages necessary for interactions with the OMDb API/data cleaning and 
manipulation from the OMDb API:

1. **tidyverse**: A collection of R packages for data science,arguably most 
useful for data manipulation and cleaning.

2. **jsonlite**: A JSON parser and generator, optimized for statistical data 
and the web.  Notably useful for interacting with a web API.

3. **imager**: Useful for loading and plotting images.

4. **ggplot2**: Useful for creation of various plots and images.

### API Interaction Functions

#### Highest Grossing Movie with **HighGross**

This function takes a vector of several Movie titles and determines which movie
was the highest grossing, along with the year of release.  It produces a
data frame ranking the movies from highest to lowest grossing.
```{r,echo=TRUE,warning=FALSE}
## Example vector of movies
m <- c("Titanic","Iron Man","Avengers: Endgame","Dances with Wolves")

## This is a supporting function to the user-input function below
Gross <- function(Title){
    Title <- gsub(" ", "+", Title)
    url <- paste("https://www.omdbapi.com/?t=",Title,"&apikey=82725ca8",sep="")
    data <- GET(url)
    # converting raw data to character
    data = rawToChar(data$content)
    # converting character to json format
    jsondata = fromJSON(data)
    # Putting news title gross earnings, and year into R object
    tgy <- c(jsondata$Title,as.numeric(parse_number(jsondata$BoxOffice)),
    jsondata$Year)
    return(tgy)
}

## Function for user-input
HighGross <- function(Movies){
  x <- data.frame(matrix(NA,nrow=length(Movies),ncol=3))
  for(i in 1:length(Movies)){
      
     x[i,] <- Gross(Movies[i])
  }
  x <- x[order(x$X2,na.last=TRUE,decreasing=TRUE),]
  colnames(x)[1] <- 'Title'
  colnames(x)[2] <- 'Gross Box Office ($)'
  colnames(x)[3] <- 'Year'
  return(x)
  }

## Example using function
HighestGrossing <- HighGross(m)
HighestGrossing
```

#### Comparing Metrics of Ratings with **HighestRating**

This function is used to explore the various online movie-rating systems, and 
the differences in ranking between each system for a given movie.  This 
function returns a vector including movie title, Internet Movie Database rating, Rotten Tomatoes rating, Metacritic rating, and IMDb rating.  These are
sorted by highest average score (standardized to a scale out of 100 for each
rating system).
```{r, echo=TRUE}
## Example vector of movies
m <- c("Titanic","Avengers: Endgame","Gone with the Wind", 
       "One Direction: This Is Us")

## This is a supporting function to the user-input function below
Ratings <- function(Title){
    Title <- gsub(" ", "+", Title)
    url <- paste("https://www.omdbapi.com/?t=",Title,"&apikey=82725ca8",sep="")
    data <- GET(url)
    # converting raw data to character
    data = rawToChar(data$content)
    # converting character to json format
    jsondata = fromJSON(data)
    # Putting news title gross earnings, and year into R object
    tgy <- c(jsondata$Title,jsondata$Ratings[1,2],jsondata$Ratings[2,2],
             jsondata$Metascore,jsondata$imdbRating)
    mean <- mean(as.numeric(substr(tgy[2],1,3))*10,
              as.numeric(substr(tgy[3],1,2)), as.numeric(tgy[4]),
              as.numeric(tgy[5])*10)
    tgy <- c(tgy, mean)
    return(tgy)
}

## User-input function
HighestRating <- function(Movies){
  x <- data.frame(matrix(NA,nrow=length(Movies),ncol=6))
  for(i in 1:length(Movies)){
      
     x[i,] <- Ratings(Movies[i])
  }
  x <- x[order(x$X6,na.last=TRUE,decreasing=TRUE),]
  colnames(x)[1] <- 'Title'
  colnames(x)[2] <- 'Internet Movie Database'
  colnames(x)[3] <- 'Rotten Tomatoes'
  colnames(x)[4] <- 'Metacritic'
  colnames(x)[5] <- 'IMDb Rating'
  colnames(x)[6] <- 'Average Rating'
  return(x)
  }
  
## Example function call
Ratings <- HighestRating(m)
Ratings

```

#### Runtime by Genre with **Runtime**

This function is used to explore the average run-time by genre (between movies presented by the user, it is not intended to be representative of all movies).
```{r, echo=TRUE}
## Example vector of movies
m <- c("Titanic","Avengers: Endgame","Gone with the Wind", 
       "One Direction: This Is Us")

## This is a supporting function to the user-input function below
RuntimeD <- function(Title){
    Title <- gsub(" ", "+", Title)
    url <- paste("https://www.omdbapi.com/?t=",Title,"&apikey=82725ca8",sep="")
    data <- GET(url)
    # converting raw data to character
    data = rawToChar(data$content)
    # converting character to json format
    jsondata = fromJSON(data)
    # Putting news title gross earnings, and year into R object
    tgy <- c(jsondata$Title,as.numeric(substr(jsondata$Runtime,1,3)),
             jsondata$Genre)
    return(tgy)
}

## User-input function
Runtimes <- function(Movies){
  x <- data.frame(matrix(NA,nrow=length(Movies),ncol=3))
  for(i in 1:length(Movies)){
      
     x[i,] <- RuntimeD(Movies[i])
  }
  x <- x[order(as.numeric(x$X2),na.last=TRUE,decreasing=TRUE),]
  colnames(x)[1] <- 'Title'
  colnames(x)[2] <- 'Runtime'
  colnames(x)[3] <- 'Genre(s)'
  return(x)
  }
  
## Example function call
Runtime <- Runtimes(m)
Runtime
```

#### Viewing Movie Posters with **poster**

This function retrieves the movie poster for a given movie and plots it with 
the help of r package **imager**.

```{r, echo=TRUE}
## Example movie vector
m <- "The Hunger Games: Mockingjay Part 1"

## This is a supporting function to the user-input function below
posterD <- function(Title){
    Title <- gsub(" ", "+", Title)
    url <- paste("https://www.omdbapi.com/?t=",Title,"&apikey=82725ca8",sep="")
    data <- GET(url)
    # converting raw data to character
    data = rawToChar(data$content)
    # converting character to json format
    jsondata = fromJSON(data)
    # Putting news title gross earnings, and year into R object
    tgy <- jsondata$Poster
    return(tgy)
}

## User-input function
Posters <- function(Movies){
  im <- load.image(posterD(Movies))
  plot(im,axes=FALSE)

}
 
## Example Function Call 
Poster <- Posters(m)
```

#### Oldest Movies (by Release Year) with **Release**

This function takes an input vector of movies and orders them by release year. 
It also outputs the title and director of each movie.

```{r, echo=TRUE}
## Example vector of movies
m <- c("Titanic","Avengers: Endgame","Gone with the Wind", 
       "One Direction: This Is Us")

## This is a supporting function to the user-input function below
ReleaseDate <- function(Title){
    Title <- gsub(" ", "+", Title)
    url <- paste("https://www.omdbapi.com/?t=",Title,"&apikey=82725ca8",sep="")
    data <- GET(url)
    # converting raw data to character
    data = rawToChar(data$content)
    # converting character to json format
    jsondata = fromJSON(data)
    # Putting news title gross earnings, and year into R object
    tgy <- c(jsondata$Title,as.numeric(jsondata$Year),jsondata$Director)
    return(tgy)
}

## User-Input function
Release <- function(Movies){
  x <- data.frame(matrix(NA,nrow=length(Movies),ncol=3))
  for(i in 1:length(Movies)){
      
     x[i,] <- ReleaseDate(Movies[i])
  }
  x <- x[order(as.numeric(x$X2),na.last=TRUE,decreasing=TRUE),]
  colnames(x)[1] <- 'Title'
  colnames(x)[2] <- 'Release Year'
  colnames(x)[3] <- 'Director'
  return(x)
  }
  
## Example Function Call
ReleaseD <- Release(m)
ReleaseD
```


#### Movies by Genre and Rating (G, PG, PG-13, R, etc.) with **GenreRate**

```{r, echo=TRUE}
## Example vectors of movies
m <- c("Titanic","Avengers: Endgame","Gone with the Wind", 
       "One Direction: This Is Us")

## This is a supporting function to the user-input function below
RateD <- function(Title){
    Title <- gsub(" ", "+", Title)
    url <- paste("https://www.omdbapi.com/?t=",Title,"&apikey=82725ca8",sep="")
    data <- GET(url)
    # converting raw data to character
    data = rawToChar(data$content)
    # converting character to json format
    jsondata = fromJSON(data)
    # Putting news title gross earnings, and year into R object
    tgy <- c(jsondata$Title,jsondata$Rated,jsondata$Genre)
    return(tgy)
}

## User-input function
Rate <- function(Movies){
  x <- data.frame(matrix(NA,nrow=length(Movies),ncol=3))
  for(i in 1:length(Movies)){
      
     x[i,] <- RateD(Movies[i])
  }
  x <- x
  colnames(x)[1] <- 'Title'
  colnames(x)[2] <- 'Rated'
  colnames(x)[3] <- 'Genre(s)'
  return(x)
  }
  
## Example Function Call
Rates <- Rate(m)
Rates
```

### Data Manipulation

Now that we have access to all of this data from the movies API, it may be 
interesting to take a look at some simple data analysis using the functions
previously created.

First, let's create a plot of movies by how well they did in the box office 
using function **HighGross**.
```{r, echo=TRUE}
# First, create a vector of many movie titles
mov <- c("Harry Potter and the Chamber of Secrets","Up","The Little Mermaid",
         "The Hunger Games","Dances with Wolves","Matilda",
         "Rogue One: A Star Wars Story","Pitch Perfect","Hook",
         "Full Metal Jacket","La La Land","Whiplash","The Jungle Book",
         "Into the Woods","The Patriot","Saving Private Ryan","The Notebook",
         "Split","The Conjuring","American Pie","Friday the 13th",
         "Avatar","Step Brothers","Happy Gilmore","The Waterboy",
         "The Big Lebowski","John Wick","The Breakfast Club","Charlotte's Web",
         "Gone with the Wind","Avengers: Endgame","Titanic","Finding Nemo",
         "Harry Potter and the Deathly Hallows: Part 2","Dune","Mamma Mia!",
         "The Goonies","Home Alone","Independence Day","Kung Fu Panda",
         "Holes")

# Call function
GrossMov <- HighGross(mov)
colnames(GrossMov)[2] <- 'y'
GrossMov <- as.data.frame(GrossMov)
GrossMov$y <- as.numeric(GrossMov$y)
## Create Plot based on this data
ggplot(data=GrossMov,aes(x = Title, y = y)) + geom_bar(stat="identity",
    fill='steelblue')+ theme_bw()+
  theme(axis.text.x=element_text(size=5,angle=45)) +
  scale_y_continuous(name = "Gross Box Office",
                     breaks=c(100000000,200000000, 300000000, 400000000,
                          500000000,600000000,700000000,800000000,850000000)) +
  ggtitle('Gross Box Office ($) By Movie')

```
This plot shows that some movies are far more successful in the box office than
others.  This leads me to wonder if there are factors contributing to the
success of these movies, such as genre.

I will explore this by retrieving the genre for the list of movies using 
**Rate** and match this with the box office gross.  I will then plot the a box
plot by genre.


```{r, echo=TRUE}
## Retrieving Movie Genres
Genre <- Rate(mov)
Genre

## Retrieving Box Office Gross
GrossMov <- HighGross(mov)
GrossMov

## Matching this data together
Table1 <- merge(GrossMov,Genre,by.x='Title',by.y='Title')
head(Table1)

## I will now classify each movie into a single genre, based on the first 
## genre listed for each movie, to simplify things.
Table1$PrimaryGenre <- word(Table1[,5],1,sep=',')
colnames(Table1)[2] <- 'Gross'
Table1$Gross <- as.numeric(Table1$Gross)

## Now, want to group these observations by genre and plot box plots of 
## the box office earnings.
ggplot(Table1, aes(x=as.factor(PrimaryGenre), y=Gross, fill=as.factor(PrimaryGenre))) + geom_boxplot() + 
    xlab("Primary Genre") + ylab("Gross Box Office ($)") +
    theme(legend.position = 'none') + ggtitle('Gross Box Office ($) By Genre')



```

Looking at this visual, it is interesting to see the differences between the 
primary genres.  Obviously, we have given the function a small sample size to 
work with, and additionally there may be outliers present that are skewing the data.  It is still interesting to see that the results are what you may expect,
with blockbuster action/advernture movies topping the box office, and animation
movies (potentially with big companies such as Disney), also seeing relatively
large box office incomes.  This data does not tell us anything statisically,
but does show an interesting exploratory analysis.

Next, I found myself interested in the "efficiency" of each movie.  That is,
I was interested in which movies made the most money per minute of movie (run-time).  To explore this, I will create a new variable that is a function
of variables from datasets taken using my previously created functions, that 
will be representative of the efficency/profitability of each movie (dollars earned per minute run time).
```{r, echo=TRUE}
## pulling runtimes
runs <- Runtimes(mov)
head(runs)
## Combining this data with previously generated table 1 that includes gross
table2 <- merge(Table1,runs,by.x="Title",by.y="Title")
table2[,-8]

## Creating a new variables in this data frame that is a function of other
## variables present
table2$productivity <- as.numeric(table2$Gross)/as.numeric(table2$Runtime)

## viewing the data
head(table2)

## Now I am going to plot productivity by runtime and see if there is any trend
## there# Change the point size, and shape
ggplot(table2, aes(x=as.numeric(Runtime),y=as.numeric(productivity),
                   color=as.factor(PrimaryGenre))) + geom_point(size=2)+
        xlab('Runtime (minutes)')+ ylab('Productivity ($/min)') +
        ggtitle('Productivity by Runtime') + 
        scale_color_discrete(name="Primary Genre")+ stat_smooth(method = "lm",
        col = "#C42126",
        se = FALSE,
        linewidth = 1)
```

Taking a look at this plot, we can see that there does not seem to be much
of a relationship between run time and productivity (at least not linearly).
To me, it speculatively seems like the movies that have the highest 
productivity are more often in the middle range of run-time.



### Wrap-Up
