---
title: "Interacting with API's - Example with the Open Movie Database API"
author: "Grace Holliday"
date: "2023-06-11"
output: github_document
  toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('httr')
library('jsonlite')
library('tidyverse')
library('imager')
```

This document is a vignette designed to introduce retrieving and manipulating
data from an API (Application Programming Interface). I will be interacting 
with the Open Movie Database API (OMDb).  I will build several functions 
comprised of a six endpoints and demonstrate some basic data manipulation. 

Endpoints include: Highest Grossing Movie, Highest Average Movie Rating,
Runtime by Genre, Viewing Movie Posters, Oldest Movies, and Movies by Genre
and Rating (PG, PG-13, etc.)

### Requirements

Packages necessary for interactions with the OMDb API/data cleaning and 
manipulation from the OMDb API:

1. **tidyverse**: A collection of R packages for data science,arguably most 
useful for data manipulation and cleaning.

2. **jsonlite**: A JSON parser and generator, optimized for statistical data 
and the web.  Notably useful for interacting with a web API.

3. **imager**: Useful for loading and plotting images.

### API Interaction Functions

#### Highest Grossing Movie with **HighGross**

This function takes a vector of several Movie titles and determines which movie
was the highest grossing, along with the year of release.  It produces a
data frame ranking the movies from highest to lowest grossing.
```{r,echo=TRUE,warning=FALSE}
## Example vector of movies
m <- c("Titanic","Iron Man","Avengers: Endgame","Dances with Wolves")

## This is a supporting function to the user-input function below
Gross <- function(Title){
    Title <- gsub(" ", "+", Title)
    url <- paste("https://www.omdbapi.com/?t=",Title,"&apikey=82725ca8",sep="")
    data <- GET(url)
    # converting raw data to character
    data = rawToChar(data$content)
    # converting character to json format
    jsondata = fromJSON(data)
    # Putting news title gross earnings, and year into R object
    tgy <- c(jsondata$Title,jsondata$BoxOffice,jsondata$Year)
    return(tgy)
}

## Function for user-input
HighGross <- function(Movies){
  x <- data.frame(matrix(NA,nrow=length(Movies),ncol=3))
  for(i in 1:length(Movies)){
      
     x[i,] <- Gross(Movies[i])
  }
  x <- x[order(x$X2,na.last=TRUE,decreasing=TRUE),]
  return(x)
  }

## Example using function
HighestGrossing <- HighGross(m)
HighestGrossing
```

#### Comparing Metrics of Ratings with **HighestRating**

This function is used to explore the various online movie-rating systems, and 
the differences in ranking between each system for a given movie.  This 
function returns a vector including movie title, Internet Movie Database rating, Rotten Tomatoes rating, Metacritic rating, and IMDb rating.  These are
sorted by highest average score (standardized to a scale out of 100 for each
rating system).
```{r, echo=TRUE}
## Example vector of movies
m <- c("Titanic","Avengers: Endgame","Gone with the Wind", 
       "One Direction: This Is Us")

## This is a supporting function to the user-input function below
Ratings <- function(Title){
    Title <- gsub(" ", "+", Title)
    url <- paste("https://www.omdbapi.com/?t=",Title,"&apikey=82725ca8",sep="")
    data <- GET(url)
    # converting raw data to character
    data = rawToChar(data$content)
    # converting character to json format
    jsondata = fromJSON(data)
    # Putting news title gross earnings, and year into R object
    tgy <- c(jsondata$Title,jsondata$Ratings[1,2],jsondata$Ratings[2,2],
             jsondata$Metascore,jsondata$imdbRating)
    mean <- mean(as.numeric(substr(tgy[2],1,3))*10,
              as.numeric(substr(tgy[3],1,2)), as.numeric(tgy[4]),
              as.numeric(tgy[5])*10)
    tgy <- c(tgy, mean)
    return(tgy)
}

## User-input function
HighestRating <- function(Movies){
  x <- data.frame(matrix(NA,nrow=length(Movies),ncol=6))
  for(i in 1:length(Movies)){
      
     x[i,] <- Ratings(Movies[i])
  }
  x <- x[order(x$X6,na.last=TRUE,decreasing=TRUE),]
  colnames(x)[1] <- 'Title'
  colnames(x)[2] <- 'Internet Movie Database'
  colnames(x)[3] <- 'Rotten Tomatoes'
  colnames(x)[4] <- 'Metacritic'
  colnames(x)[5] <- 'IMDb Rating'
  colnames(x)[6] <- 'Average Rating'
  return(x)
  }
  
## Example function call
Ratings <- HighestRating(m)
Ratings

```

#### Runtime by Genre with **Runtime**

This function is used to explore the average run-time by genre (between movies presented by the user, it is not intended to be representative of all movies).
```{r, echo=TRUE}
## Example vector of movies
m <- c("Titanic","Avengers: Endgame","Gone with the Wind", 
       "One Direction: This Is Us")

## This is a supporting function to the user-input function below
RuntimeD <- function(Title){
    Title <- gsub(" ", "+", Title)
    url <- paste("https://www.omdbapi.com/?t=",Title,"&apikey=82725ca8",sep="")
    data <- GET(url)
    # converting raw data to character
    data = rawToChar(data$content)
    # converting character to json format
    jsondata = fromJSON(data)
    # Putting news title gross earnings, and year into R object
    tgy <- c(jsondata$Title,as.numeric(substr(jsondata$Runtime,1,3)),
             jsondata$Genre)
    return(tgy)
}

## User-input function
Runtimes <- function(Movies){
  x <- data.frame(matrix(NA,nrow=length(Movies),ncol=3))
  for(i in 1:length(Movies)){
      
     x[i,] <- RuntimeD(Movies[i])
  }
  x <- x[order(as.numeric(x$X2),na.last=TRUE,decreasing=TRUE),]
  colnames(x)[1] <- 'Title'
  colnames(x)[2] <- 'Runtime'
  colnames(x)[3] <- 'Genre(s)'
  return(x)
  }
  
## Example function call
Runtime <- Runtimes(m)
Runtime
```

#### Viewing Movie Posters with **poster**

This function retrieves the movie poster for a given movie and plots it with 
the help of r package **imager**.

```{r, echo=TRUE}
## Example movie vector
m <- "The Hunger Games: Mockingjay Part 1"

## This is a supporting function to the user-input function below
posterD <- function(Title){
    Title <- gsub(" ", "+", Title)
    url <- paste("https://www.omdbapi.com/?t=",Title,"&apikey=82725ca8",sep="")
    data <- GET(url)
    # converting raw data to character
    data = rawToChar(data$content)
    # converting character to json format
    jsondata = fromJSON(data)
    # Putting news title gross earnings, and year into R object
    tgy <- jsondata$Poster
    return(tgy)
}

## User-input function
Posters <- function(Movies){
  im <- load.image(posterD(Movies))
  plot(im,axes=FALSE)

}
 
## Example Function Call 
Poster <- Posters(m)
```

#### Oldest Movies (by Release Year) with **Release**

This function takes an input vector of movies and orders them by release year. 
It also outputs the title and director of each movie.

```{r, echo=TRUE}
## Example vector of movies
m <- c("Titanic","Avengers: Endgame","Gone with the Wind", 
       "One Direction: This Is Us")

## This is a supporting function to the user-input function below
ReleaseDate <- function(Title){
    Title <- gsub(" ", "+", Title)
    url <- paste("https://www.omdbapi.com/?t=",Title,"&apikey=82725ca8",sep="")
    data <- GET(url)
    # converting raw data to character
    data = rawToChar(data$content)
    # converting character to json format
    jsondata = fromJSON(data)
    # Putting news title gross earnings, and year into R object
    tgy <- c(jsondata$Title,as.numeric(jsondata$Year),jsondata$Director)
    return(tgy)
}

## User-Input function
Release <- function(Movies){
  x <- data.frame(matrix(NA,nrow=length(Movies),ncol=3))
  for(i in 1:length(Movies)){
      
     x[i,] <- ReleaseDate(Movies[i])
  }
  x <- x[order(as.numeric(x$X2),na.last=TRUE,decreasing=TRUE),]
  colnames(x)[1] <- 'Title'
  colnames(x)[2] <- 'Release Year'
  colnames(x)[3] <- 'Director'
  return(x)
  }
  
## Example Function Call
ReleaseD <- Release(m)
ReleaseD
```


#### Movies by Genre and Rating (G, PG, PG-13, R, etc.) with **GenreRate**

```{r, echo=TRUE}
## Example vectors of movies
m <- c("Titanic","Avengers: Endgame","Gone with the Wind", 
       "One Direction: This Is Us")

## This is a supporting function to the user-input function below
RateD <- function(Title){
    Title <- gsub(" ", "+", Title)
    url <- paste("https://www.omdbapi.com/?t=",Title,"&apikey=82725ca8",sep="")
    data <- GET(url)
    # converting raw data to character
    data = rawToChar(data$content)
    # converting character to json format
    jsondata = fromJSON(data)
    # Putting news title gross earnings, and year into R object
    tgy <- c(jsondata$Title,jsondata$Rated,jsondata$Genre)
    return(tgy)
}

## User-input function
Rate <- function(Movies){
  x <- data.frame(matrix(NA,nrow=length(Movies),ncol=3))
  for(i in 1:length(Movies)){
      
     x[i,] <- RateD(Movies[i])
  }
  x <- x
  colnames(x)[1] <- 'Title'
  colnames(x)[2] <- 'Rated'
  colnames(x)[3] <- 'Genre(s)'
  return(x)
  }
  
## Example Function Call
Rates <- Rate(m)
Rates
```

### Data Manipulation


### Wrap-Up
